---
author: 3ikakke
title: '14. Statistics Visualizations'
duration: 50
footer: '3ikkke Data Science with Python'
output: slidy_presentation
date: 'Friday 8th of April 2022'
---

## Outline  
- Objectives  
- Review of probabilities  
- Introducing two-way tables  
- Joint probabilities    
- Conditional probabilities    
- Review of Distributions   
- Univariate numeric analysis  
- Univariate categorical analysis  
- Bivariate: numeric vs numeric  
- Bivariate: categorical vs categorical  
- Bivariate: numeric vs categorical  

## Objectives  
- Reinforce data exploration  
- Understand the value of 2 way tables  
- Learn data vizualization rules based on univariate and bivariate analysis  


## Review of probabilities  
- Remember the rules?  
	+ Probabilities exist between 0 and 1  
	+ All probabilities in a system add up to 1  
	+ The compliment rule: if there are only 2 probabilities then 1 - known probability = unknown probability  
	+ $P(A\cup{B}) = P(A) + P(B) - P(A\cap{B})$ : remember the double counting problem  
- Joint probabilities: $P(A\cap{B})$ means both conditions are true  
- Conditional probabilities: $P(A|B)$ means probability of A being true only in the group of B (read as Probability of A given B)  

## Introducing two-way tables  
*Distribution of Pass and Fail among Private and Public Schools*
```{r echo=F, message=F, warning=F}
library(knitr)
a <- as.data.frame(matrix(c(23, 12, 17, 44), nrow=2))
row.names(a) = c('Passed', 'Failed')
names(a) = c('Private', 'Public')
kable(a)
```  

- Let's use keys to identify <span style='color: red'>Private</span> as **R**, <span style='color: red'>Public</span> as **U**, <span style='color: red'>Pass</span> as **A** and <span style='color: red'>Fail</span> as **F**  
- Total = 23 + 12 + 17 + 44 = **96**  
- Probability of attending a private school or <span style='color: red'>$P(R)$</span> = (23 + 12)/96 = 0.3646 = **36.46%**  
- Probability of attending a public school or <span style='color: red'>$P(U)$</span> = (17 + 44)/96 = 0.6354 = **63.54%**  
- Probability of passing or <span style='color: red'>$P(A)$</span> = (23+17)/96 = 0.4167 = **41.67%**  
- Probability of failing or <span style='color: red'>$P(F)$</span> = (12+44)/96 = 0.5833 = **58.33%**  

## Two-way tables: Joint probabilities    

```{r echo=F, message=F, warning=F}
kable(a)
```  
- Probability of attending a private school and passing (joint probability) is <span style='color: red'>$P(R\cap{A})$</span> = 23/96 = 0.2396 = **23.96%**  
- Probability of attending a private school and failing (joint prbability) is <span style='color: red'>$P(R\cap{F})$</span> = 12/96 = 0.1250 = **12.50%**  
- Probability of attending a public school and passing (joint probability) is <span style='color: red'>$P(U\cap{A})$</span> = 17/96 = 0.1771 = **17.71%**  
- Probability of attending a pubic school and failing (joint probability) is <span style='color: red'>$P(U\cap{F})$</span> = 44/ = 0.4583 = **45.83%**  
- All probabilities are between 0 and 1  
- Add all joint probabilities and they come to 1! **0.2396 + 0.1771 + 0.1250 + 0.4583  = 1** 


## Two-way tables: Conditional probabilities    
```{r echo=F, message=F, warning=F}
kable(a)
```  
- Probability of a person passing if they attend a private school (conditional probability) is <span style='color: red'>$P(A|R)$</span> = 23/(23+12) = **0.66**  
- Probability of a person failing if they attend a private school (conditional probability) is <span style='color: red'>$P(F|R)$</span> = 12/(23+12) = **0.34**  
- Probability of a person passing if they attend a public school (conditional probability) is <span style='color: red'>$P(A|U)$</span> = 17/(17+44) = **0.28**  
- Probability of a person failing if they attend a public school (conditional probability) is <span style='color: red'>$P(F|U)$</span> = 44/(44+17) = **0.72**  
- Probability that a person attended a private school if they passed (conditional probability) is <span style='color: red'>$P(R|P)$</span> = 23/(23+17) = **0.58**  
- Probability that a person attended a public school if they passed (conditional probability) is <span style='color: red'>$P(U|P)$</span> = 17/(23+17) = **0.43**  
- Probability that a person attended a private school if they failed (conditional probability) is <span style='color: red'>$P(R|F)$</span> = 12/(12+44) = **0.21**  
- Probability that a person attended a public school if they failed (conditional probability) is <span style='color: red'>$P(U|F)$</span> = 44/(12+44) = **0.79**     
- **How would these help you make a decision?**  


## Review of Distributions   
- The histogram  
	+ A bell curve is a normal distribution    
	+ The mean approximates the median in a normal distribution   
	+ Report median and IQR for non-normal distributions  
	+ Report mean and standard deviation for normal distributions  
- The five number summary:
	+ Minimum  
	+ First Quartile (Q1)  
	+ Median  
	+ Third Quartile (Q3)  
	+ Maximum  
- The skeleton of a boxplot  

## Univariate numeric analysis  
- We need matplotlib and seaborn  
```python
import pandas as pd  
from matplotlib import pyplot as plt  
import seaborn as sns  

#lets read in our data  
dataset = pd.read_csv("nhanes.csv")
```  
- Histograms  
- Boxplots  
```python 

sns.histplot(x='num_var', data=dataset)
plt.show() 

sns.boxplot(x='num_var', data=dataset)
plt.show() 
```  


## Univariate categorical analyis  
- Frequencies  
- Proportions  
- Percentages  
- Barplots (count plots)    
- Pie charts are a bad idea!  
```python  
sns.countplot(x='cat_var', data=dataset)
```  

## Bivariate: numeric vs numeric  
- Correlation (Pearson correlation coefficient)  
	+ -1 to 0 to +1 
	+ -1 indicates negative coorelation (the numbers trend in opposite directions)  
	+ 0 means no correlation  
	+ +1 means the numbers trend in the same direction  
- Scatterplots 
```python  
sns.scatterplot(x='num_var1', y='num_var2', data=dataset)
```  

## Bivariate: categorical vs categorical  
- Two way tables  
- Grouped barplots  
```python 
sns.countplot(x='cat_var1', hue='cat_var2', data=dataset)
```  


## Bivariate: numeric vs categorical  
- Rememeber - report the numeric summaries at each level of the categorical variable  
```python 
sns.boxplot(x='num_var', y='cat_var', data=dataset)
```  

## Q&A  

## Review of objectives  
- Reinforce data exploration  
- Understand the value of 2 way tables  
- Learn data vizualization rules based on univariate and bivariate analysis  

## Gist of the day  
- Get the pdf from [here](./14.Statistics.Visuals.pdf)  
- Get the [gist]()    
- The Jupyter Notebook will be uploaded  

## Thanks for being awesome students!  


